The state learner \\ {-- a super learner for right-censored data}

### 2024-03-21 tex/jrss-version/state-learner-jrss-style.tex
changed notation to -1 for censoring
change k to j in appendix

%% Toggle with or without figures
\def\nofig{1}

@@

The problem with most existing super learners for right-censored data is that
they depend on a pre-specified estimator of the censoring distribution. The main
idea of the state learner is to jointly use learners of \( \Lambda_1 \),
\( \Lambda_2 \), and \( \Gamma \), and the relations in
equation~(\ref{eq:parametrizeP}), to learn a feature of the observed data
distribution \( P \). The discrete state learner ranks a tuple of learners of
\( (\Lambda_1, \Lambda_2, \Gamma) \) based on how well they jointly model the
observed data. To formally introduce the state learner, we define the
multi-state process
\begin{equation*}
  \eta(t) = \1\{\tilde{T} \leq t, \tilde D=1\} + 2\,\1\{\tilde{T} \leq t, \tilde D=2\}+ 3\,\1\{\tilde{T} \leq t, \tilde D=0\},
  \quad \text{for} \quad t \in [0, \tau].
\end{equation*}
Note that at time \(t\), we observe that each observation is in one of four
mutually exclusive states (Figure \ref{fig:multi-state-process}).
\begin{figure}[h]
  \centering %
  \includeFigCond[width=.5\textwidth]{figure-multi-state-process.pdf}
  \caption{Illustration of the multi-state process \(\eta\) used by
    the state learner. Note that `censored' is a state, hence the
    process is always observed at any time.}
  \label{fig:multi-state-process}
\end{figure}
The conditional distribution of \( \eta(t) \) given \( X=x \) is determined by
the \fxnote*{reconsider letting censoring state be \( -1 \)?}{function}
\begin{equation}
  \label{eq:F-def}
  F(t, k, x) = P(\eta(t) = k \mid X=x),
  \quad \text{for all} \quad
  t \in [0,\tau],
  k \in \{0,1,2,3\},
  x \in \mathcal{X}.
\end{equation}
The function \( F \) describes the conditional state occupation
probabilities corresponding to the observed multi-state process
\(\eta\).

@@

  F(t, 3, x)

@@

    = 1- F(t, 1, x) - F(t, 2, x)- F(t, 3, x).

@@

  \phi_{a_1,a_2, b}(\data_n)(t,3,x) &= \int_0^t e^{\{-a_1(\data_n)(s \mid x)-a_2(\data_n)(s \mid x) - b(\data_n)(s \mid x)\} }  b(\data_n)(\diff s \mid x),\\
  \phi_{a_1,a_2, b}(\data_n)(t,0,x) &= 1-  \sum_{j=1}^3\phi_{a_1,a_2, b}(\data_n)(t,j,x).

@@

  B_t(F,O) = \sum_{j=0}^{3}

@@

\begin{proposition}
  \label{prop:stric-prop}
  For \(P_0\in\mathcal{P}\) define
  \begin{equation*}
    F^* = \argmin_{F} P_0{[\bar{B}_\tau(F, \blank)]},
  % F^* = \argmin_{(\Lambda_1,\Lambda_2,\Gamma) \in \mathcal{M}\otimes\mathcal{M}\otimes\mathcal{M}} P_0{[\bar B_\tau(\phi_{(\Lambda_1,\Lambda_2,\Gamma)}, \blank)]},
\end{equation*}
where the minimum is taken over all \( F \), such that \( F \) is a conditional
state occupation probability function for some measure \( P \) as defined in equation~(\ref{eq:F-def}).
Then \( F^*(t,j,\blank) = F_0(t,j,\blank) \) \( H \)-almost surely for
any \( j\in \{0,1,2,3\} \) and almost any \( t \in [0, \tau]\).
\end{proposition}
\begin{proof}
  See Appendix~\ref{sec:proof-proposition}.
\end{proof}

We establish a finite sample oracle result for the state learner.
Our Corollary~\ref{cor:oracle-prop} is in essence a special case of a
general cross-validation result by \cite{vaart2006oracle}. We assume
that we split the data into equally sized folds, and for simplicity of
presentation we take \( n \) to be such that
\( |\data_n^{-k}| = n/K \) with \( K \) fixed. We will allow the
number of learners to grow with \( n \) and write
\( \mathcal{F}_n=\mathcal{F}(\mathcal{A}_{1,n}, \mathcal{A}_{2,n},
\mathcal{B}_n)\) as short-hand notation and to emphasise the
dependence on \( n \).
% The discrete super $\hat{\phi}_n$ and the
% oracle $\tilde{\phi}_n$ are defined as in Section~\ref{sec:framework}
% but now with respect to the library \( \mathcal{F}_n \) and the loss
% \( B \).
In the following we let \( \| \blank \|_{P_0} \) denote the norm
\begin{equation}
  \label{eq:norm}
  \| F \|_{P_0} = 
  \left\{
    \sum_{j=0}^{3}\int_{\mathcal{X}} \int_0^{\tau} F(t, j, x)^2 \diff t H_0( \diff x)
  \right\}^{1/2}.
\end{equation}
\begin{corollary}
  \label{cor:oracle-prop}
  For all \(P_0\in\mathcal{P}\), \( n \in \N \), \( k \in \{1, \dots, K\} \),
  and $\delta>0$,
  \begin{align*}
    \E_{P_0}{\left[ \Vert \hat{\phi}_n(\data_n^{-k}) - F_0 \Vert_{P_0}^2 \right]}
    & \leq (1 + 2\delta)
      \E_{P_0}{\left[ \Vert \tilde{\phi}_n(\data_n^{-k}) - F_0 \Vert_{P_0}^2 \right]}
    \\
    & \quad
      + (1+ \delta) 16   K \tau
      \left(
      13 + \frac{12}{\delta}
      \right)
      \frac{\log(1 + |\mathcal{F}_n|)}{n}.
  \end{align*}
\end{corollary}
\begin{proof}
  See Appendix~\ref{sec:proof-proposition}.
\end{proof}

@@

In Section~\ref{sec:targeted-learning} we demonstrate how the state learner can
be used for targeted learning. A targeted estimator is typically obtained from
estimators of the nuisance parameters $\Lambda_1$, $\Lambda_2$, and $\Gamma$. By
equations~(\ref{eq:lambdaj}) and~(\ref{eq:gamma}) and the definition of \( F \),
we have
\begin{equation}
  \label{eq:7}
  \Gamma(t , x) 
  = \int_0^t  \frac{F(\diff s, 3, x )}{F(s-, 0, x )},
  \quad \text{and} \quad
  \Lambda_j(t , x) 
  = \int_0^t  \frac{F(\diff s, j, x )}{F(s-, 0, x )},
  \quad j \in \{1,2\},
\end{equation}
and thus a targeted estimator can also be obtained from an estimator of $F$
using equation~(\ref{eq:7}). The key feature of a targeted estimator is that it
is asymptotically equivalent to a sum of i.i.d.\ random variables plus a second
order remainder term \citep{bickel1993efficient,fisher2021visually}. For our
setting of competing risks, the remainder term is typically dominated by terms
of the form
\begin{equation}
  \label{eq:dr-term}
  P{\left[
      \int_0^{\tau} w_n(s, \blank)
      \hat{M}_{1,n}(s \mid  \blank)
      \hat{M}_{2,n}(\diff s \mid  \blank)
    \right]},
\end{equation}
where \( (\hat{M}_{1,n}, \hat{M}_{2,n}) \) is any of the nine combinations of
\( \hat{M}_{1,n} \in \{[\Gamma -\hat{\Gamma}_n], [\Lambda_1
-\hat{\Lambda}_{1,n}], [\Lambda_2 -\hat{\Lambda}_{2,n}]\} \) and
\( \hat{M}_{2,n} \in \{[\Gamma -\hat{\Gamma}_n], [\Lambda_1
-\hat{\Lambda}_{1,n}], [\Lambda_2 -\hat{\Lambda}_{2,n}]\} \), and \( w_n \) is
some data-dependent function with domain \([0,\tau]\times\mathcal X \)
\citep{van2003unified}. In particular, a targeted estimator will be
asymptotically linear if the `products' of the estimation errors
\( \hat{M}_{1,n} \) and \( \hat{M}_{2,n} \) in equation~(\ref{eq:dr-term}) are
\( \smallO_P{(n^{-1/2})}\). Proposition~\ref{prop:dr-structure} states that if
equation~(\ref{eq:dr-term}) holds for the a targeted estimator based on
estimators $\hat{\Lambda}_{1,n}$, $\hat{\Lambda}_{2,n}$, and $\hat{\Gamma}_{n}$,
then a similar product structure holds for a targeted estimator based on
\( \hat{F}_n \). We state the result for the special case that
\(\hat{M}_{1,n}= \Gamma-\hat{\Gamma}_n \) and
\(\hat{M}_{2,n} =\Lambda_1-\hat{\Lambda}_{1,n} \), but similar results holds for
any combinations of \( \Gamma-\hat{\Gamma}_n\),
\( \Lambda_1-\hat{\Lambda}_{1,n} \), and \( \Lambda_2-\hat{\Lambda}_{2,n} \).
\begin{proposition}
  \label{prop:dr-structure}
  Assume that \( w(s,x)\leq c \), \( F(s, 0, x) \geq 1/c \) and
  \( \hat{F}_n(s, 0, x) \geq 1/c \) for some \( c>0 \) for all
  \( s \in [0, \tau] \) and \( x \in \mathcal{X} \). Then there are real-valued
  uniformly bounded functions \( w^a_n \), \( w^b_n \), \( w^c_n \), and
  \( w^d_n \) with domain \( [0,\tau]^2 \times \mathcal{X} \) such that
  \begin{align*}
    & P_0{\left[
      \int_0^{\tau} w(s, \blank)
      \left\{
      \Gamma_0(s,\blank) -\hat{\Gamma}_n(s,\blank)
      \right\}
      [\Lambda_0-\hat{\Lambda}_n]
      (\diff s, \blank)
      \right]}
    \\
    & =
      P_0{\left[
      \int_0^{\tau} \int_0^{s} w^a_n(s,u,\blank) [F_0 - \hat{F}_n](u-, 0, \blank)[F_0 - \hat{F}_n](s-, 0, \blank) F_0(\diff u, 2, \blank ) F_0 ( \diff s, 1, \blank)
      \right]}
    \\
    & \quad +
      P_0{\left[
      \int_0^{\tau} \int_0^{s} w^b_n(s,u,\blank) [F_0 - \hat{F}_n](u-, 0, \blank)
      F_0(\diff u, 2, \blank ) [F_0 - \hat{F}_n](\diff s, 1, \blank)
      \right]}
    \\
    & \quad +
      P_0{\left[
      \int_0^{\tau} \int_0^{s} w^c_n(s,u,\blank) [F_0 - \hat{F}_n](\diff u, 2, \blank)
      [F_0 - \hat{F}_n](s-, 0, \blank)
      F_0(\diff s, 1, \blank ) 
      \right]}
    \\
    & \quad +
      P_0{\left[
      \int_0^{\tau} \int_0^{s} w^d_n(s,u,\blank) [F_0 - \hat{F}_n](\diff u, 2, \blank)
      [F_0 - \hat{F}_n](\diff s, 1, \blank)
      \right]}.
  \end{align*}
\end{proposition}
\begin{proof}
  See Appendix~\ref{sec:state-learner-with}.
\end{proof}

@@

Define
\( \bar{B}_{\tau,0}(F, o) = \bar{B}_{\tau}(F, o) - \bar{B}_{\tau}(F_0, o) \) and
\( R_{0}(F) = P_0{[\bar{B}_{\tau,0}(F, \blank)]} \).
\begin{lemma}
  \label{lemma:norm}
  \( R_{0}(F) = \Vert F - F_0 \Vert_{P_0}^2 \), where \( \Vert \blank \Vert_{P_0}\) is defined
  in equation~(\ref{eq:norm}).
\end{lemma}
\begin{proof}
  For any \( t \in [0, \tau] \) and \( k\in \{0,1,2\} \) we have
  \begin{align*}
    & \E_{P_0}{\left[ (F(t, k, X) - \1{\{\eta(t) = k \}})^2 \right]}
    \\
    & =    \E_{P_0}{\left[ (F(t, k, X) - F_0(t, k, X) + F_0(t, k, X) - \1{\{\eta(t) = k
      \}})^2 \right]}
    \\
    & =    \E_{P_0}{\left[ (F(t, k, X) - F_0(t, k, X))^2\right]}
      + \E_{P_0}{\left[ (F_0(t, k, X) - \1{\{\eta(t) = k \}})^2\right]}
    \\
    & \quad
      + 2\E_{P_0}{\left[ (F(t, k, X) - F_0(t, k, X))(F_0(t, k, X) - \1{\{\eta(t) = k
      \}})\right]}
    \\
    & =    \E_{P_0}{\left[ (F(t, k, X) - F_0(t, k, X))^2\right]}
      + \E_{P_0}{\left[ (F_0(t, k, X) - \1{\{\eta(t) = k \}})^2\right]},
  \end{align*}
  where the last equality follows from the tower property. Hence, using Fubini,
  we have
  \begin{equation*}
    P{[\bar{B}_{\tau}(F, \blank)]}
    = \Vert F - F_0 \Vert_{P_0}^2 + P_0{[\bar{B}_{\tau}(F_0, \blank)]}.
  \end{equation*}
\end{proof}

>>>>

%% Toggle with or without figures
% \def\nofig{1}

@@

The problem with most existing super learners for right-censored data is that
they depend on a pre-specified estimator of the censoring distribution. The main
idea of the state learner is to jointly use learners of \( \Lambda_1 \),
\( \Lambda_2 \), and \( \Gamma \), and the relations in
equation~(\ref{eq:parametrizeP}), to learn a feature of the observed data
distribution \( P \). The discrete state learner ranks a tuple of learners of
\( (\Lambda_1, \Lambda_2, \Gamma) \) based on how well they jointly model the
observed data. To formally introduce the state learner, we define the
multi-state process
\begin{equation*}
  \eta(t) = \1\{\tilde{T} \leq t, \tilde D=1\} + 2\,\1\{\tilde{T} \leq t, \tilde
  D=2\} - \1\{\tilde{T} \leq t, \tilde D=0\},
  \quad \text{for} \quad t \in [0, \tau].
\end{equation*}
Note that at time \(t\), we observe that each observation is in one of four
mutually exclusive states (Figure \ref{fig:multi-state-process}).
\begin{figure}[h]
  \centering %
  \includeFigCond[width=.5\textwidth]{figure-multi-state-process.pdf}
  \caption{Illustration of the multi-state process \(\eta\) used by
    the state learner. Note that `censored' is a state, hence the
    process is always observed at any time.}
  \label{fig:multi-state-process}
\end{figure}
The conditional distribution of \( \eta(t) \) given \( X=x \) is determined by
the function
\begin{equation}
  \label{eq:F-def}
  F(t, k, x) = P(\eta(t) = k \mid X=x),
  \quad \text{for all} \quad
  t \in [0,\tau],
  k \in \{-1,0,1,2\},
  x \in \mathcal{X}.
\end{equation}
The function \( F \) describes the conditional state occupation
probabilities corresponding to the observed multi-state process
\(\eta\).

@@

  F(t, -1, x)

@@

    = 1- F(t, 1, x) - F(t, 2, x)- F(t, -1, x).

@@

  \phi_{a_1,a_2, b}(\data_n)(t,-1,x) &= \int_0^t e^{\{-a_1(\data_n)(s \mid x)-a_2(\data_n)(s \mid x) - b(\data_n)(s \mid x)\} }  b(\data_n)(\diff s \mid x),\\
  \phi_{a_1,a_2, b}(\data_n)(t,0,x) &= 1
                                      - \phi_{a_1,a_2, b}(\data_n)(t,1,x)
                                      - \phi_{a_1,a_2, b}(\data_n)(t,2,x)
  & \qquad
    -\phi_{a_1,a_2, b}(\data_n)(t,-1,x)

@@

  B_t(F,O) = \sum_{j=-1}^{2}

@@

\begin{proposition}
  \label{prop:stric-prop}
  For \(P_0\in\mathcal{P}\) define
  \begin{equation*}
    F^* = \argmin_{F} P_0{[\bar{B}_\tau(F, \blank)]},
  % F^* = \argmin_{(\Lambda_1,\Lambda_2,\Gamma) \in \mathcal{M}\otimes\mathcal{M}\otimes\mathcal{M}} P_0{[\bar B_\tau(\phi_{(\Lambda_1,\Lambda_2,\Gamma)}, \blank)]},
\end{equation*}
where the minimum is taken over all \( F \), such that \( F \) is a conditional
state occupation probability function for some measure \( P \) as defined in equation~(\ref{eq:F-def}).
Then \( F^*(t,j,\blank) = F_0(t,j,\blank) \) \( H \)-almost surely for
any \( j\in \{-1,0,1,2\} \) and almost any \( t \in [0, \tau]\).
\end{proposition}
\begin{proof}
  See Appendix~\ref{sec:proof-proposition}.
\end{proof}

We establish a finite sample oracle result for the state learner.
Our Corollary~\ref{cor:oracle-prop} is in essence a special case of a
general cross-validation result by \cite{vaart2006oracle}. We assume
that we split the data into equally sized folds, and for simplicity of
presentation we take \( n \) to be such that
\( |\data_n^{-k}| = n/K \) with \( K \) fixed. We will allow the
number of learners to grow with \( n \) and write
\( \mathcal{F}_n=\mathcal{F}(\mathcal{A}_{1,n}, \mathcal{A}_{2,n},
\mathcal{B}_n)\) as short-hand notation and to emphasise the
dependence on \( n \).
% The discrete super $\hat{\phi}_n$ and the
% oracle $\tilde{\phi}_n$ are defined as in Section~\ref{sec:framework}
% but now with respect to the library \( \mathcal{F}_n \) and the loss
% \( B \).
In the following we let \( \| \blank \|_{P_0} \) denote the norm
\begin{equation}
  \label{eq:norm}
  \| F \|_{P_0} = 
  \left\{
    \sum_{j=-1}^{2}\int_{\mathcal{X}} \int_0^{\tau} F(t, j, x)^2 \diff t H_0( \diff x)
  \right\}^{1/2}.
\end{equation}
\begin{corollary}
  \label{cor:oracle-prop}
  For all \(P_0\in\mathcal{P}\), \( n \in \N \), \( k \in \{1, \dots, K\} \),
  and $\delta>0$,
  \begin{align*}
    \E_{P_0}{\left[ \Vert \hat{\phi}_n(\data_n^{-k}) - F_0 \Vert_{P_0}^2 \right]}
    & \leq (1 + 2\delta)
      \E_{P_0}{\left[ \Vert \tilde{\phi}_n(\data_n^{-k}) - F_0 \Vert_{P_0}^2 \right]}
    \\
    & \quad
      + (1+ \delta) 16   K \tau
      \left(
      13 + \frac{12}{\delta}
      \right)
      \frac{\log(1 + |\mathcal{F}_n|)}{n}.
  \end{align*}
\end{corollary}
\begin{proof}
  See Appendix~\ref{sec:proof-proposition}.
\end{proof}

@@

In Section~\ref{sec:targeted-learning} we demonstrate how the state learner can
be used for targeted learning. A targeted estimator is typically obtained from
estimators of the nuisance parameters $\Lambda_1$, $\Lambda_2$, and $\Gamma$. By
equations~(\ref{eq:lambdaj}) and~(\ref{eq:gamma}) and the definition of \( F \),
we have
\begin{equation}
  \label{eq:7}
  \Gamma(t , x) 
  = \int_0^t  \frac{F(\diff s, -1, x )}{F(s-, 0, x )},
  \quad \text{and} \quad
  \Lambda_j(t , x) 
  = \int_0^t  \frac{F(\diff s, j, x )}{F(s-, 0, x )},
  \quad j \in \{1,2\},
\end{equation}
and thus a targeted estimator can also be obtained from an estimator of $F$
using equation~(\ref{eq:7}). The key feature of a targeted estimator is that it
is asymptotically equivalent to a sum of i.i.d.\ random variables plus a second
order remainder term \citep{bickel1993efficient,fisher2021visually}. For our
setting of competing risks, the remainder term is typically dominated by terms
of the form
\begin{equation}
  \label{eq:dr-term}
  P{\left[
      \int_0^{\tau} w_n(s, \blank)
      \hat{M}_{1,n}(s \mid  \blank)
      \hat{M}_{2,n}(\diff s \mid  \blank)
    \right]},
\end{equation}
where \( (\hat{M}_{1,n}, \hat{M}_{2,n}) \) is any of the nine combinations of
\( \hat{M}_{1,n} \in \{[\Gamma -\hat{\Gamma}_n], [\Lambda_1
-\hat{\Lambda}_{1,n}], [\Lambda_2 -\hat{\Lambda}_{2,n}]\} \) and
\( \hat{M}_{2,n} \in \{[\Gamma -\hat{\Gamma}_n], [\Lambda_1
-\hat{\Lambda}_{1,n}], [\Lambda_2 -\hat{\Lambda}_{2,n}]\} \), and \( w_n \) is
some data-dependent function with domain \([0,\tau]\times\mathcal X \)
\citep{van2003unified}. In particular, a targeted estimator will be
asymptotically linear if the `products' of the estimation errors
\( \hat{M}_{1,n} \) and \( \hat{M}_{2,n} \) in equation~(\ref{eq:dr-term}) are
\( \smallO_P{(n^{-1/2})}\). Proposition~\ref{prop:dr-structure} states that if
equation~(\ref{eq:dr-term}) holds for the a targeted estimator based on
estimators $\hat{\Lambda}_{1,n}$, $\hat{\Lambda}_{2,n}$, and $\hat{\Gamma}_{n}$,
then a similar product structure holds for a targeted estimator based on
\( \hat{F}_n \). We state the result for the special case that
\(\hat{M}_{1,n}= \Gamma-\hat{\Gamma}_n \) and
\(\hat{M}_{2,n} =\Lambda_1-\hat{\Lambda}_{1,n} \), but similar results holds for
any combinations of \( \Gamma-\hat{\Gamma}_n\),
\( \Lambda_1-\hat{\Lambda}_{1,n} \), and \( \Lambda_2-\hat{\Lambda}_{2,n} \).
\begin{proposition}
  \label{prop:dr-structure}
  Assume that \( w(s,x)\leq c \), \( F(s, 0, x) \geq 1/c \) and
  \( \hat{F}_n(s, 0, x) \geq 1/c \) for some \( c>0 \) for all
  \( s \in [0, \tau] \) and \( x \in \mathcal{X} \). Then there are real-valued
  uniformly bounded functions \( w^a_n \), \( w^b_n \), \( w^c_n \), and
  \( w^d_n \) with domain \( [0,\tau]^2 \times \mathcal{X} \) such that
  \begin{align*}
    & P_0{\left[
      \int_0^{\tau} w(s, \blank)
      \left\{
      \Gamma_0(s,\blank) -\hat{\Gamma}_n(s,\blank)
      \right\}
      [\Lambda_0-\hat{\Lambda}_n]
      (\diff s, \blank)
      \right]}
    \\
    & =
      P_0{\left[
      \int_0^{\tau} \int_0^{s} w^a_n(s,u,\blank) [F_0 - \hat{F}_n](u-, 0, \blank)[F_0 - \hat{F}_n](s-, 0, \blank) F_0(\diff u, 2, \blank ) F_0 ( \diff s, 1, \blank)
      \right]}
    \\
    & \quad +
      P_0{\left[
      \int_0^{\tau} \int_0^{s} w^b_n(s,u,\blank) [F_0 - \hat{F}_n](u-, 0, \blank)
      F_0(\diff u, 2, \blank ) [F_0 - \hat{F}_n](\diff s, 1, \blank)
      \right]}
    \\
    & \quad +
      P_0{\left[
      \int_0^{\tau} \int_0^{s} w^c_n(s,u,\blank) [F_0 - \hat{F}_n](\diff u, 2, \blank)
      [F_0 - \hat{F}_n](s-, 0, \blank)
      F_0(\diff s, 1, \blank ) 
      \right]}
    \\
    & \quad +
      P_0{\left[
      \int_0^{\tau} \int_0^{s} w^d_n(s,u,\blank) [F_0 - \hat{F}_n](\diff u, 2, \blank)
      [F_0 - \hat{F}_n](\diff s, 1, \blank)
      \right]}.
  \end{align*}
\end{proposition}
\begin{proof}
  See Appendix~\ref{sec:state-learner-with}.
\end{proof}

@@

Define
\( \bar{B}_{\tau,0}(F, o) = \bar{B}_{\tau}(F, o) - \bar{B}_{\tau}(F_0, o) \) and
\( R_{0}(F) = P_0{[\bar{B}_{\tau,0}(F, \blank)]} \).
\begin{lemma}
  \label{lemma:norm}
  \( R_{0}(F) = \Vert F - F_0 \Vert_{P_0}^2 \), where \( \Vert \blank \Vert_{P_0}\) is defined
  in equation~(\ref{eq:norm}).
\end{lemma}
\begin{proof}
  For any \( t \in [0, \tau] \) and \( j\in \{-1,0,1,2\} \) we have
  \begin{align*}
    & \E_{P_0}{\left[ (F(t, j, X) - \1{\{\eta(t) = j \}})^2 \right]}
    \\
    & =    \E_{P_0}{\left[ (F(t, j, X) - F_0(t, j, X) + F_0(t, j, X) - \1{\{\eta(t) = j
      \}})^2 \right]}
    \\
    & =    \E_{P_0}{\left[ (F(t, j, X) - F_0(t, j, X))^2\right]}
      + \E_{P_0}{\left[ (F_0(t, j, X) - \1{\{\eta(t) = j \}})^2\right]}
    \\
    & \quad
      + 2\E_{P_0}{\left[ (F(t, j, X) - F_0(t, j, X))(F_0(t, j, X) - \1{\{\eta(t) = j
      \}})\right]}
    \\
    & =    \E_{P_0}{\left[ (F(t, j, X) - F_0(t, j, X))^2\right]}
      + \E_{P_0}{\left[ (F_0(t, j, X) - \1{\{\eta(t) = j \}})^2\right]},
  \end{align*}
  where the last equality follows from the tower property. Hence, using Fubini,
  we have
  \begin{equation*}
    P{[\bar{B}_{\tau}(F, \blank)]}
    = \Vert F - F_0 \Vert_{P_0}^2 + P_0{[\bar{B}_{\tau}(F_0, \blank)]}.
  \end{equation*}
\end{proof}
